### 2019-04-08 papers/journal/encoding.tex
Todos done
Minor Fixes

The following constraint encodes no conditions on activities on nodes,
i.e., all the present molecules on the nodes are active.
\begin{align*}
\texttt{FunNoN} =& \bigwedge\limits_{i,m} n_{i,m} = a_{i,m}  
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the node is
defined by a Boolean function $\nodef_m$ of presence of molecules present on that node.
\begin{align*}
\texttt{FunBfN} =& \bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
\end{align*}
The following constraint encodes no conditions on activities on edges,
i.e., all the present molecules on the edges are active.
\begin{align*}
 \texttt{FunNoE} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m}
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the edge is defined by a Boolean function $\edgef_m$ of the presence of molecules present on that edge.
\begin{align*}
 \texttt{FunBfE} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_k(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
\end{align*}
%
The following constraint encodes that the activity of the molecules on
edges is defined by inhibition by other molecules based on the pairing
matrix. 
\begin{align*}
\texttt{FunPairE} =&	\begin{aligned}[t]
\bigwedge\limits_{i,j,q, m}  [e_{i,j,q,m} \land \bigwedge_{m_l \neq m,l \in \{1,..,3\}} e_{i,j,q,m'} \land p_{m, m_l}] \limplies\\
	(\neg b_{i,j,q,m} \land \bigwedge_{\mathclap{m_l \neq m,l, p_{m, m_l} \in \{1,..,3\}}} \neg b_{i,j,q,m_l})\end{aligned}\\
\end{align*}
%
We denote the particular choice of node and edge function combined by \texttt{ActivityFun}.
\begin{align*}
\texttt{ActivityFun} = \texttt{NodeFun}_m \land \texttt{EdgeFun}_m
\end{align*}
\subsubsection{Constraints for stability condition}
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following constraint recursively encodes that node $j$ is $m$-reachable from node $i$ in less than $p$ steps
if either there is a direct edge between $i$ and $j$ with $m$ present on the edge or there is a edge between $i^{\prime \prime}$ and
$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
\begin{align*}
  \texttt{Paths}(r) &= \bigwedge\limits_{i,j,m,p} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
\end{align*}
Now we can encode stability using the reachability variables.
and say if there is an $m$-edge between $i$ and $j$, there is
$m$-reachable path from $j$ to $i$.
\begin{align*}
 \texttt{Loop}(r) &= \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies r_{j,i,m,\nu}
\end{align*}
We define \texttt{StabilityC} to collect all the stability related constraints.
% 
\texttt{StabilityC} states that every molecule is moving in a cycle.
\begin{align*}
\texttt{StabilityC} = \texttt{Paths}(r) \land \texttt{Loop}(r) 
\end{align*}
% Boolean variable
% F3 and F2 are used to model these constraints.

@@

To specify the stability condition, we use \improve{Nondet usage is not clear in the encoding. improve.}nondeterminism to perform enumeration.
%
%Nondeterminism can be viewed as an oracle that guesses a correct path for a specific $m$ if it exists. 
%
%We then use this path as a witness to ensures that stability is maintained for each outgoing molecule $m$ on every edge $E$.
%
%
Note that the enumeration is done only for an unwinding depth of $w$.
%
The encoding presented below is a simpler version of the implementation. 
\subsubsection{Constraints for stability condition}
%
A path between node $i$ and $j$ in the graph is an $m$-$path_{ij}$ if molecule $m$ is present on edges along that path.
%
%We use {\npath} to represent the required path to encode the stability condition in VTS.
%
%A naive implementation is shown in~\ref{eq:reachbmc1} where 
We use {\npath}$(m, i, j, w)$ to represent the presence of a $m$-$path_{ij}$ of maximum $w$ length.
%there is a edge between $i^{\prime \prime}$ and01
%$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
% \begin{align}
%   NOracle (m, i, j, w) = \exists \, path[h1, ..., hn]: \\ 2 \leq |path| \leq  w + 2  \, \land h1 = i \, \land  hn = j \, \land \\ 
%   \bigvee\limits_{k = 1}^{|path|- 1} \bigvee\limits_{q}  E[hk] [hk+1][q][m] 
%   \tag{S1}\label{eq:reach1}
% \end{align}
%The NOracle can be thought of as a disjunction of $m$-path of increment length. 
%
$\npath(m, i, j, w)$ performs enumeration of every $m$-$path_{ij}$ of at most $w$ length. 
%
Starting with exploring $m$-${paths}_{ij}$ of length 1; a direct edge between $i$ and $j$ with molecule $m$ present, until every $m$-${paths}_{ij}$ of length $w-1$ is explored.\change{m-path looks - sign in the align}
%by performing series of node hops till final node destination j is reached.
%Starting with $m$-${paths}_i^j$ of length 1 which is a direct edge between $i$ and $j$ with molecule $m$ present on the edge or performing series of node hops till final node destination j is reached.

%|m-path| = 1 \lor  |m-path| = 2 \lor ...  |m-path| = w.     \\
%\bigvee\limits_{l = 1}^{|path| - 1} |m-path| = l

\begin{align*}
\npath(m, i, j, w) = \bigvee\limits_{l = 1}^{w-1} (\texttt{Length} (m-path_{ij}) == l)
\end{align*}

>>>>

The following constraint encodes no conditions on activities on nodes,
i.e., all the present molecules on the nodes are active.
\begin{align*}
\texttt{FunNoN} =& \bigwedge\limits_{i,m} n_{i,m} = a_{i,m}  
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the node is
defined by a Boolean function $\nodef_m$ of presence of molecules present on that node.
\begin{align*}
\texttt{FunBfN} =& \bigwedge\limits_{i,m} n_{i,m} \limplies a_{i,m} =  \nodef_m (n_{i,1},\dots,n_{i,\mu}) 
\end{align*}
The following constraint encodes no conditions on activities on edges,
i.e., all the present molecules on the edges are active.
\begin{align*}
 \texttt{FunNoE} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} = b_{i,j,q,m}
\end{align*}
The following constraint encodes that activity of a molecule $m$ on the edge is defined by a Boolean function $\edgef_m$ of the presence of molecules present on that edge.
\begin{align*}
 \texttt{FunBfE} =& \bigwedge\limits_{i,j,q,m} e_{i,j,q,m} \limplies b_{i,j,q,m} = \edgef_k(e_{i,j,q,1}, .., e_{i,j,q,\mu} )
\end{align*}
%
The following constraint encodes that the activity of the molecules on
edges is defined by inhibition by other molecules based on the pairing
matrix. 
\begin{align*}
\texttt{FunPairE} =&	\begin{aligned}[t]
\bigwedge\limits_{i,j,q, m}  [e_{i,j,q,m} \land \bigwedge_{m_l \neq m,l \in \{1,..,3\}} e_{i,j,q,m'} \land p_{m, m_l}] \limplies\\
	(\neg b_{i,j,q,m} \land \bigwedge_{\mathclap{m_l \neq m,l, p_{m, m_l} \in \{1,..,3\}}} \neg b_{i,j,q,m_l})\end{aligned}\\
\end{align*}
%
We denote the particular choice of node and edge function combined by \texttt{ActivityFun}.
\begin{align*}
\texttt{ActivityFun} = \texttt{NodeFun}_m \land \texttt{EdgeFun}_m
\end{align*}
\subsubsection{Constraints for stability condition}
%
We use $m$-reachability to encode the stability condition in VTSs.
%
The following constraint recursively encodes that node $j$ is $m$-reachable from node $i$ in less than $p$ steps
if either there is a direct edge between $i$ and $j$ with $m$ present on the edge or there is a edge between $i^{\prime \prime}$ and
$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
\begin{align*}
  \texttt{Paths}(r) &= \bigwedge\limits_{i,j,m,p} r_{i,j,m,p} \limplies (\bigvee_{q} \, e_{i,j,q,m} \lor \bigvee_{i\neq i^{\prime}} ( \, \bigvee_{q} e_{i,i^{\prime},q,m}) \land r_{i^{\prime},j,m,p-1} )
\end{align*}
Now we can encode stability using the reachability variables.
and say if there is an $m$-edge between $i$ and $j$, there is
$m$-reachable path from $j$ to $i$.
\begin{align*}
 \texttt{Loop}(r) &= \bigwedge\limits_{i,j,m} (\bigvee_{q} e_{i,j,q,m}) \limplies r_{j,i,m,\nu}
\end{align*}
We define \texttt{StabilityC} to group all the stability related constraints.
% 
\texttt{StabilityC} states that every molecule is moving in a cycle.
\begin{align*}
\texttt{StabilityC} = \texttt{Paths}(r) \land \texttt{Loop}(r) 
\end{align*}
% Boolean variable
% F3 and F2 are used to model these constraints.

@@

To specify the stability condition, we use nondeterminism to perform enumeration.
%
%Nondeterminism can be viewed as an oracle that guesses a correct path for a specific $m$ if it exists. 
%
%We then use this path as a witness to ensures that stability is maintained for each outgoing molecule $m$ on every edge $E$.
%
%
Note that the enumeration is done only for an unwinding depth of $w$.
%
The encoding presented below is a simpler version of the implementation. 
\subsubsection{Constraints for stability condition}
%
A path between node $i$ and $j$ in the graph is an $m$-$path_{ij}$ if molecule $m$ is present on edges along that path.
%
%We use {\npath} to represent the required path to encode the stability condition in VTS.
%
%A naive implementation is shown in~\ref{eq:reachbmc1} where 
We use {\npath}$(m, i, j, w)$ to represent the presence of a $m$-$path_{ij}$ of maximum $w$ length.
%there is a edge between $i^{\prime \prime}$ and01
%$(i \neq i^{\prime \prime})$ with $m$ present, and j is $m$-reachable from $i'$ in less than $p-1$ steps.
%
% \begin{align}
%   NOracle (m, i, j, w) = \exists \, path[h1, ..., hn]: \\ 2 \leq |path| \leq  w + 2  \, \land h1 = i \, \land  hn = j \, \land \\ 
%   \bigvee\limits_{k = 1}^{|path|- 1} \bigvee\limits_{q}  E[hk] [hk+1][q][m] 
%   \tag{S1}\label{eq:reach1}
% \end{align}
%The NOracle can be thought of as a disjunction of $m$-path of increment length. 
%
$\npath(m, i, j, w)$ performs enumeration of every $m$-$path_{ij}$ of at most $w$ length. 
%
Starting with exploring $m$-${paths}_{ij}$ of length 1; a direct edge between $i$ and $j$ with molecule $m$ present, until every $m$-${paths}_{ij}$ of length $w-1$ is explored.
%by performing series of node hops till final node destination j is reached.
%Starting with $m$-${paths}_i^j$ of length 1 which is a direct edge between $i$ and $j$ with molecule $m$ present on the edge or performing series of node hops till final node destination j is reached.
%
%|m-path| = 1 \lor  |m-path| = 2 \lor ...  |m-path| = w.     \\
%\bigvee\limits_{l = 1}^{|path| - 1} |m-path| = l
%
\begin{align*}
\npath(m, i, j, w) = \bigvee\limits_{l = 1}^{w-1} (\texttt{Length} (m-path_{ij}) == l)
\end{align*}

### 2019-04-08 papers/journal/intro.tex
Minor Fixes

\subsubsection{Vesicle budding} \unsure{Somya should we say it involve three mojor steps? This line is not optimal.}The process of vesicle budding involves sorting and packaging of cargo molecules, membrane deformation and budding of the vesicle from the source compartment (Figure~\ref{fig:vts}).
%
Cargo sorting is performed by cytosolic adaptor proteins. 
Adaptors are recruited to the source compartment through interactions with the lipids that form the compartment's membrane, and interactions with molecular switches called Arf/Sar which signal the initiation of vesicle formation~\cite{paczkowski2015cargo}. 
%
Then these adaptors bind to cargo molecules and thereby sequester them for packaging. 
%
Different adaptor proteins act at different compartments, for example, vesicle formation at the ER requires the adaptor sec23/24, and the AP2 adaptor is used for packaging vesicles at the plasma membrane.
%

@@

One such benchmark is provided by our result about the connectivity of the VTS graph. Graph connectivity is defined as the minimum number of edges that need to be removed from an undirected graph to render it disconnected. For example, in a directed cyclic graph, every node has at least one incoming and one outgoing edge, at least 2 edges need to be removed in order to disconnect the graph. Therefore the underlying undirected graph is 2-connected. 
From our result, for a given set of biomolecular interactions, for a cell that is in a steady state, we expect the VTS graph to have a certain k-connectivity. When regulation is given by arbitrary Boolean rules, we find that VTS graphs must be 3-connected. And when we add the constraint of the specific form of regulatory interactions seen between VTS molecules, we expect vesicle transport graphs to be \change{We have used 4 connected in results! 2 strongly or 4 what to put?}2-strongly connected.

>>>>

\subsubsection{Vesicle budding} The process of vesicle budding involves three major steps, sorting and packaging of cargo molecules, membrane deformation and budding of the vesicle from the source compartment (Figure~\ref{fig:vts}).
%
Cargo sorting is performed by cytosolic adaptor proteins. 
Adaptors are recruited to the source compartment through interactions with the lipids that form the compartment's membrane, and interactions with molecular switches called Arf/Sar which signal the initiation of vesicle formation~\cite{paczkowski2015cargo}. 
%
Then these adaptors bind to cargo molecules and thereby sequester them for packaging. 
%
Different adaptor proteins act at different compartments, for example, vesicle formation at the ER requires the adaptor sec23/24, and the AP2 adaptor is used for packaging vesicles at the plasma membrane.
%

@@

One such benchmark is provided by our result about the connectivity of the VTS graph. Graph connectivity is defined as the minimum number of edges that need to be removed from an undirected graph to render it disconnected. For example, in a directed cyclic graph, every node has at least one incoming and one outgoing edge, at least 2 edges need to be removed in order to disconnect the graph. Therefore the underlying undirected graph is 2-connected. 
From our result, for a given set of biomolecular interactions, for a cell that is in a steady state, we expect the VTS graph to have a certain k-connectivity. When regulation is given by arbitrary Boolean rules, we find that VTS graphs must be 3-connected. And when we add the constraint of the specific form of regulatory interactions seen between VTS molecules, we expect vesicle transport graphs to be 4-connected.
